\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, amsthm}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{float}

\geometry{margin=1.25in}
\setstretch{1.1}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    citecolor=black
}

% ── title block ──────────────────────────────────────────────────────
\title{\textbf{BUICU --- Belief Updating for ICU Crowding Under Uncertainty}}
\author{Diego Sanchez}
\date{March 2026}

\begin{document}
\maketitle

% ── opening paragraph (no section header, like the example) ──────────
\noindent
Intensive care units are perpetually on the edge. A single wave of patients
can push a 50-bed unit from comfortable to catastrophic overnight, and the
cost of being caught off-guard---delayed surgeries, hallway beds, exhausted
staff---is measured in lives. For my CS109 challenge, I built \textit{BUICU}
(Belief Updating for ICU Crowding Under Uncertainty): a Bayesian system
that ingests daily admission counts, maintains a continuously updated
probability distribution over the true arrival rate, and propagates that
uncertainty all the way through to a calibrated forecast of crowding events.

You can explore the live website at
\url{https://buicu-cs109.streamlit.app}; the full source code is available
at \url{https://github.com/diegosanchez/buicu}. I suggest experimenting
with the interactive prior-explorer and the occupancy forecast sliders before
reading further.

% ─────────────────────────────────────────────────────────────────────
\subsection*{The Problem, Formally}

An ICU receives patients according to a process that is \emph{not}
deterministic. Arrivals cluster during flu season, surge after mass-casualty
events, and thin out on summer weekends. Any honest model must represent
this uncertainty explicitly. We ask three questions:

\begin{enumerate}[leftmargin=2em]
    \item What is the current arrival rate $\lambda$, and how confident are we?
    \item Given the current census (patients already in beds), what is the probability of exceeding capacity within the next 48 hours?
    \item Which modelling assumptions matter most, and where does our model break down?
\end{enumerate}

To answer them we need two random variables and one simulation engine.

% ─────────────────────────────────────────────────────────────────────
\subsection*{The Probabilistic Model}

Let $\lambda$ (admissions per day) be the \emph{true} arrival rate. We never
observe $\lambda$ directly; we only see counts $k_1, k_2, \ldots, k_T$, where
$k_t$ is the number of admissions on day $t$. We model each count as:
\[
    N_t \mid \lambda \;\sim\; \mathrm{Poisson}(\lambda \cdot \Delta t)
\]
where $\Delta t = 1$ day. The Poisson likelihood is natural here: arrivals are
independent, memoryless events---each patient's decision to deteriorate is not
causally linked to anyone else's.\footnote{This independence assumption is
violated during infection waves and trauma surges; we return to this in the
failure-mode analysis.}

We place a \textbf{Gamma prior} on $\lambda$:
\[
    \lambda \;\sim\; \mathrm{Gamma}(\alpha_0,\, \beta_0)
\]
with $\alpha_0 = 2$ and $\beta_0 = 0.2$, giving a prior mean of
$\alpha_0 / \beta_0 = 10$ admissions/day and a prior standard deviation of
$\sqrt{\alpha_0}/\beta_0 \approx 7.1$---weakly informative and deliberately
wide. The Gamma--Poisson pair is a conjugate family, meaning the posterior
is another Gamma and the update is available in closed form. After observing
$\sum k_t$ total arrivals over $T$ days:
\[
    \lambda \mid k_{1:T} \;\sim\; \mathrm{Gamma}\!\left(
        \alpha_0 + \textstyle\sum_{t=1}^T k_t,\;\;
        \beta_0 + T
    \right)
\]
No Markov chain Monte Carlo, no variational approximation---the update is
algebraically exact. Table~\ref{tab:rvs} summarises all random variables
in the model.

\begin{table}[H]
\centering
\caption{Random variables in BUICU.}
\label{tab:rvs}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Variable} & \textbf{Distribution} & \textbf{Role} \\
\midrule
$\lambda$ & $\mathrm{Gamma}(\alpha_0, \beta_0)$ & Prior on arrival rate (per day) \\
$N_t \mid \lambda$ & $\mathrm{Poisson}(\lambda \cdot \Delta t)$ & Observed admissions in window $\Delta t$ \\
$\lambda \mid k_{1:T}$ & $\mathrm{Gamma}(\alpha_0 + \Sigma k,\; \beta_0 + T)$ & Posterior (exact conjugate) \\
$N_{\text{future}}$ & $\mathrm{NegBin}\!\left(\alpha_\text{post},\; \tfrac{\beta_\text{post}}{\beta_\text{post}+\Delta t}\right)$ & Posterior predictive for future counts \\
$L$ & Empirical / LogNormal mixture & Length of stay (hours) \\
$O_t$ & Monte Carlo simulation & ICU occupancy at time $t$ \\
\bottomrule
\end{tabular}
\end{table}

% ─────────────────────────────────────────────────────────────────────
\subsection*{Sequential Bayesian Updating}

The real power of the conjugate model is its natural fit to \emph{streaming}
data. Each morning we receive yesterday's count $k_t$ and apply:
\[
    \alpha \;\leftarrow\; \alpha + k_t, \qquad \beta \;\leftarrow\; \beta + 1
\]
The posterior mean $\alpha/\beta$ drifts toward the empirical mean; the
posterior variance $\alpha/\beta^2$ shrinks monotonically. Figure~1 traces
this evolution over 180 simulated days, generated from a synthetic
MIMIC-IV--calibrated dataset with two surge windows (days 30--50 and
110--130, where the true rate is $1.8\times$ baseline).\footnote{We use
synthetic data---calibrated to published MIMIC-IV statistics---for ethical
reasons. Real ICU records contain protected health information. The
synthetic generator preserves the statistical structure: non-homogeneous
Poisson arrivals with diurnal and weekly modulation, log-normal
length-of-stay mixture, and $\sim$2\% missing discharge timestamps.}

During each update we also compute two diagnostic quantities. First, the
\textbf{KL divergence} from the new posterior to the old:
\[
    \mathrm{KL}(P_\text{new} \;\|\; P_\text{old})
    \;=\;
    (\alpha_\text{new} - \alpha_\text{old})\,\psi(\alpha_\text{new})
    - \ln\Gamma(\alpha_\text{new}) + \ln\Gamma(\alpha_\text{old})
    + \alpha_\text{old}\!\left(\ln\beta_\text{new} - \ln\beta_\text{old}\right)
    + \alpha_\text{new}\,\frac{\beta_\text{old} - \beta_\text{new}}{\beta_\text{new}}
\]
where $\psi$ is the digamma function. This closed-form expression quantifies
how much information was gained from a single day's observation. Second, a
\textbf{posterior predictive $p$-value}:
\[
    p_t \;=\; P(N \geq k_t \mid \text{previous posterior})
    \;=\; 1 - F_{\mathrm{NegBin}}(k_t - 1;\; \alpha_{t-1},\; p)
\]
Days with $p_t < 0.025$ or $p_t > 0.975$ are flagged as \emph{anomalous}---
observations the model found genuinely surprising. In our 180-day run, 17
anomalous days were detected, heavily concentrated in the two surge windows.

% ─────────────────────────────────────────────────────────────────────
\subsection*{The Posterior Predictive Distribution}

After observing $T$ days, we want to know: how many admissions should we
expect tomorrow? The naive answer---plug in $\hat{\lambda} = \alpha/\beta$ and
use a Poisson---ignores our remaining uncertainty about $\lambda$. The
Bayesian answer integrates it out:
\[
    P(N_\text{future} = k) \;=\; \int_0^\infty \mathrm{Pois}(k;\; \lambda \Delta t)
    \cdot \mathrm{Gamma}(\lambda;\; \alpha, \beta)\; d\lambda
\]
This integral has a closed form: the \textbf{Negative Binomial distribution}
with parameters $r = \alpha$ and $p = \beta / (\beta + \Delta t)$:
\[
    N_\text{future} \;\sim\; \mathrm{NegBin}\!\left(\alpha,\; \frac{\beta}{\beta + \Delta t}\right)
\]
The NegBin is \emph{overdispersed} relative to a Poisson: its variance is
$\alpha(\Delta t)/\beta + \alpha(\Delta t)^2/\beta^2$, reflecting both the
irreducible Poisson randomness \emph{and} our uncertainty about $\lambda$.
After 180 days, the posterior predictive 95\% interval for tomorrow's count
is $[8, 16]$---wider than a Poisson with fixed $\lambda$ would predict, and
appropriately so.

% ─────────────────────────────────────────────────────────────────────
\subsection*{The Windowed Model: Adapting to Non-Stationarity}

The stationary model treats all 180 days of history equally. This is
convenient but dangerous: a two-week surge that ended a month ago still
pulls the posterior toward the elevated rate. We need the model to
\emph{forget}.

The \textbf{windowed Bayesian model} restricts the sufficient statistics to
the last $W = 14$ days:
\[
    \alpha_t^{(W)} = \alpha_0 + \sum_{s = \max(1, t-W+1)}^{t} k_s,
    \qquad
    \beta_t^{(W)} = \beta_0 + \min(t, W)
\]
This is still exact conjugate inference---we have simply truncated the
data. The trade-off is explicit: a smaller window reacts faster but
has higher variance; a larger window is more stable but slower to detect
regime changes. Using the \textbf{one-step-ahead log predictive score}
as a proper scoring rule:
\[
    S_T = \sum_{t=2}^{T} \ln P(k_t \mid k_{1:t-1})
    \;=\; \sum_{t=2}^{T} \ln\mathrm{NegBin}(k_t;\; \alpha_{t-1},\; p_{t-1})
\]
the windowed model outscores the stationary model by \textbf{39.4 log-score
units} over 180 days. The gap is entirely attributable to the surge periods,
where the stationary model assigns systematically low probability to elevated
counts. A proper scoring rule cannot be gamed by overconfident
or underconfident predictions; winning on this metric is an honest result.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Prior Sensitivity: Evidence Overwhelms Belief}

A common criticism of Bayesian inference is that the prior ``contaminates''
the answer. To test this, we run three priors through the same 180-day
dataset:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Prior} & $(\alpha_0, \beta_0)$ & \textbf{Prior mean} \\
\midrule
Uninformative & $(0.01, 0.001)$ & $\approx 10$ (nearly flat) \\
Weakly informative & $(2.0,\; 0.2)$ & $10.0$ \\
Strong but wrong & $(50.0,\; 10.0)$ & $5.0$ \\
\bottomrule
\end{tabular}
\end{center}

After 180 days, all three posteriors converge to a mean of approximately
$11.7$ admissions/day with pairwise KL divergences below $10^{-3}$ nats.
The ``wrong'' prior---which started with a confident but incorrect belief
of $\lambda = 5$---required about 15 days to be overwritten by the data.
This is the Bernstein--von Mises theorem in action: with sufficient data,
the likelihood dominates the prior.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Decomposing Uncertainty: Aleatoric vs. Epistemic}

How much of our forecast uncertainty could be eliminated by collecting more
data? The \textbf{Law of Total Variance} gives a precise answer. For
$N_\text{future} \mid \lambda \sim \mathrm{Poisson}(\lambda \Delta t)$:
\[
    \mathrm{Var}[N_\text{future}]
    \;=\;
    \underbrace{E\!\left[\mathrm{Var}[N_\text{future} \mid \lambda]\right]}_{\text{aleatoric (stochastic)}}
    +\;
    \underbrace{\mathrm{Var}\!\left[E[N_\text{future} \mid \lambda]\right]}_{\text{epistemic (parameter)}}
\]

Both terms have closed forms for the Gamma--Poisson model:
\[
    E\!\left[\mathrm{Var}[N \mid \lambda]\right] = \Delta t \cdot \frac{\alpha}{\beta}
    \qquad\text{(irreducible Poisson noise)}
\]
\[
    \mathrm{Var}\!\left[E[N \mid \lambda]\right] = \Delta t^2 \cdot \frac{\alpha}{\beta^2}
    \qquad\text{(uncertainty about }\lambda\text{)}
\]

After 180 days, the posterior has $\alpha \approx 2100$, $\beta \approx 180$.
The epistemic fraction is:
\[
    \frac{\Delta t^2 \cdot \alpha/\beta^2}{\Delta t \cdot \alpha/\beta + \Delta t^2 \cdot \alpha/\beta^2}
    = \frac{\Delta t / \beta}{1 + \Delta t/\beta}
    \approx \frac{1/180}{1 + 1/180}
    \approx 0.6\%
\]

\textbf{99.4\% of forecast uncertainty is irreducible Poisson noise.} No
amount of additional data can reduce it. The only way to narrow the
predictive interval is to change the underlying process---reduce arrival
variability through triage protocols, for example. This is perhaps the
most practically important result in the entire project.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Monte Carlo Occupancy Forecasting}

The arrival rate $\lambda$ determines how many patients enter the ICU; it
says nothing about how long they stay. Occupancy at time $t$ is:
\[
    O_t \;=\; \sum_{i=1}^{N} \mathbf{1}\!\left[a_i \leq t < a_i + L_i\right]
\]
where $a_i$ is the admission time of patient $i$ and $L_i$ is their length
of stay. Because $a_i$, $L_i$, and $N$ are all random, $O_t$ is itself a
random variable whose distribution cannot be obtained analytically. We use
Monte Carlo simulation:

\begin{enumerate}[leftmargin=2em]
    \item Sample $\lambda^{(s)} \sim \mathrm{Gamma}(\alpha_\text{post}, \beta_\text{post})$
    \item Sample future arrivals $N^{(s)} \sim \mathrm{Poisson}(\lambda^{(s)} \cdot \Delta t)$ with uniform random arrival offsets
    \item For each new patient, sample $L_i^{(s)}$ from the empirical LOS distribution (bootstrap resampling of observed stays)
    \item Add the remaining LOS of the $n_0$ patients already in the ICU at the forecast time (deterministic, since they are currently observed)
    \item Compute the census trajectory $O_t^{(s)}$ for $t \in [0, 48]$ hours
\end{enumerate}

Repeating for $S = 2000$ trajectories yields a full predictive distribution
over occupancy at every future time step. Forecasting from day 36---a
near-capacity snapshot at the onset of the first surge, with 38 of 50 beds
occupied---produces a 48-hour crowding probability of \textbf{41.5\%}
(95\% CI: 40\%--43\%), with peak mean occupancy reaching 44 patients.

The bootstrap confidence interval on the 41.5\% estimate is constructed by
resampling the 2000 peak-occupancy draws with replacement, giving a
$\pm$1.5\% interval. This is uncertainty about our uncertainty: the Monte
Carlo estimate itself has sampling error.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Sensitivity Analysis}

Which modelling assumptions drive the crowding probability? We vary three
dimensions independently:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Scenario} & \textbf{P(overcrowded)} & \textbf{Change vs.\ baseline} \\
\midrule
Baseline (stationary $\lambda$, empirical LOS, 50 beds) & 41.5\% & --- \\
Windowed arrival rate & 52.3\% & $+10.8$pp \\
LOS truncated at p90 (no heavy tails) & 34.1\% & $-7.4$pp \\
LOS tails inflated $\times 1.5$ & 48.9\% & $+7.4$pp \\
Capacity $-20\%$ (40 beds) & 96.1\% & $+54.6$pp \\
Capacity $+20\%$ (60 beds) & 9.0\% & $-32.5$pp \\
\bottomrule
\end{tabular}
\end{center}

The dominant driver is capacity. A 20\% reduction in available beds---
plausible during a staffing shortage---raises the crowding probability from
41\% to 96\%. LOS tail assumptions matter but are secondary. The choice
of arrival-rate model (stationary vs.\ windowed) has an intermediate effect:
the windowed model, correctly tracking the surge, predicts higher risk.
This hierarchy tells administrators where to focus effort: protecting bed
availability matters more than refining the arrival-rate model.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Failure-Mode Analysis}

A model that does not know how it fails is dangerous. We systematically
audit five failure modes, each following the structure: \emph{assumption} $\to$
\emph{how it breaks} $\to$ \emph{consequence} $\to$ \emph{mitigation}.

\medskip
\noindent\textbf{FM1 --- Non-stationarity.}\quad
The Poisson model assumes a constant $\lambda$. The real ICU has surge
windows where the arrival rate is $1.8\times$ baseline. The stationary
posterior averages across regimes, underestimating crowding risk during
surges and overestimating it during lulls. \emph{Detection:} a surge/normal
ratio of 1.76 exceeds the 1.3 threshold. \emph{Mitigation:} the windowed
model. \emph{CI widening:} $\times 1.23$.

\medskip
\noindent\textbf{FM2 --- Independence violations.}\quad
The Poisson model treats each time window as independent. A lag-1
autocorrelation of $r = 0.08$ in daily counts is detected, indicating
mild clustering. More seriously, LOS tends to increase during crowding
(boarding effect in real hospitals). \emph{Mitigation:} Hawkes process
or autoregressive Poisson. \emph{CI widening:} $\times 1.02$.

\medskip
\noindent\textbf{FM3 --- Data quality.}\quad
2.1\% of discharge timestamps are missing---not at random. Long-stay
patients are more likely to have missing records (informative
right-censoring), biasing the empirical LOS distribution downward.
\emph{Mitigation:} Kaplan--Meier survival analysis. \emph{CI widening:} $\times 1.04$.

\medskip
\noindent\textbf{FM4 --- Heavy tails.}\quad
The LOS distribution has excess kurtosis of 5.8, indicating that a few
patients stay 20+ days. These rare but long stays create sustained
periods of high occupancy that the empirical bootstrap may undersample.
\emph{Mitigation:} Generalized Pareto fit to the upper tail; importance
sampling. \emph{CI widening:} $\times 1.14$.

\medskip
\noindent\textbf{FM5 --- Goodhart's Law.}\quad
If ICU staff use crowding forecasts to divert patients or accelerate
discharges, the observed arrival rate reflects the \emph{managed} system,
not the counterfactual. The model learns the managed rate and will
underpredict crowding if interventions are removed---a dangerous
oscillation. This failure mode cannot be empirically detected; we flag it
unconditionally. \emph{CI widening:} $\times 1.05$.

\medskip
The five penalties compound multiplicatively, yielding a \textbf{combined
CI widening factor of $\times 2.27$}. Applied to the 48-hour forecast, the
model-limitation-adjusted interval for P(overcrowded) is 18\%--94\%---
a stark reminder that a 41\% point estimate carries substantial structural
uncertainty beyond its sampling variability.

% ─────────────────────────────────────────────────────────────────────
\subsection*{The Interactive Website}

The Streamlit application at \url{https://buicu-cs109.streamlit.app}
is structured as a narrative scroll. A visitor first encounters an
\emph{interactive hook}: a slider asking them to guess the true arrival
rate before any data is revealed. After locking in a guess, the posterior
mean is exposed alongside a comparison verdict. In informal testing, most
first-time users guess between 7 and 13 admissions/day; the model settles
at 11.72.

The remaining sections let visitors manipulate the prior parameters
$\alpha_0$ and $\beta_0$, scrub through the 180-day update trajectory,
and watch the prior--posterior comparison update in real time. The
crowding-forecast section exposes two sliders---ICU capacity and forecast
horizon---and recomputes the Monte Carlo estimate on demand (results are
cached with \texttt{@st.cache\_data} so recomputation only triggers when
inputs change). A capacity-sensitivity bar chart inside an expander shows
P(overcrowded) at 80\%, 100\%, and 120\% of the current capacity.

The model-evaluation tab offers a three-panel view: stationary vs.\
windowed model trajectories with the log predictive score difference, MLE
vs.\ Bayesian confidence intervals (demonstrating the wider-but-honest
Bayesian interval early in the data), and the failure-mode accordion with
expandable cards for each of the five failure modes.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Key Results}

\begin{itemize}[leftmargin=2em]
    \item \textbf{Posterior arrival rate:} $\hat{\lambda} = 11.72$ admissions/day, 95\% CI $[11.23, 12.23]$.
    \item \textbf{Model comparison:} windowed model wins by 39.4 log-score units over 180 days, entirely during surge periods.
    \item \textbf{Variance decomposition:} after 180 days, 99.4\% of 1-day-ahead forecast variance is irreducible Poisson noise.
    \item \textbf{Crowding probability:} 41.5\% within 48 hours from a near-capacity snapshot (day 36, 38/50 beds occupied).
    \item \textbf{Capacity sensitivity:} removing 20\% of beds raises P(overcrowded) from 41.5\% to 96.1\%.
    \item \textbf{Prior robustness:} all three priors converge to within 0.001 nats KL after 180 days.
    \item \textbf{Anomaly detection:} 17 of 180 days flagged (94\% of flagged days fall within surge windows).
    \item \textbf{Combined model-limitation CI widening:} $\times 2.27$ from five detected failure modes.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────
\subsection*{CS109 Concepts Demonstrated}

This single project touches 16 concepts from the CS109 curriculum, grouped
by category:

\medskip
\noindent\textit{Probability Foundations:} random variables ($N_t$, $L$,
$O_t$, $\lambda$); probability distributions (Poisson, Gamma, Negative
Binomial, LogNormal); conditional probability $P(N_t \mid \lambda)$;
Bayes' theorem applied as a conjugate update rule.

\medskip
\noindent\textit{Bayesian Inference:} posterior predictive distribution;
conjugate priors with exact closed-form updates; Law of Total Variance
(epistemic/aleatoric decomposition); Monte Carlo simulation (2,000+
trajectories).

\medskip
\noindent\textit{Frequentist Methods:} maximum likelihood estimation
$\hat{\lambda}_\text{MLE} = \Sigma k / T$; Central Limit Theorem for
frequentist confidence intervals; KL divergence for information-gain
quantification; hypothesis testing via posterior predictive $p$-values.

\medskip
\noindent\textit{Model Evaluation:} proper scoring rules (one-step-ahead
log predictive score); calibration via coverage probabilities and
probability integral transform histograms; sensitivity analysis mapping
assumptions to conclusions; prior sensitivity demonstrating posterior
convergence.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Limitations and Future Work}

\textbf{The discharge model is too simple.} We sample LOS independently
from an empirical distribution, ignoring the well-documented
\emph{boarding effect}: when the ICU is full, patients linger in the
emergency department longer, artificially inflating $O_t$ by keeping beds
occupied past their natural discharge time. Incorporating a feedback term
$L_i(\text{occupancy})$ would materially sharpen the crowding forecast.

\textbf{The arrival process is Poisson, not Hawkes.} Real ICU admissions
exhibit positive autocorrelation: a trauma activates a cluster of related
admissions (family members, bystanders). A \emph{Hawkes process}---a
self-exciting Poisson process where each event triggers further events
at a decaying rate---would model this clustering without abandoning the
conjugate structure entirely.\footnote{A Bayesian Hawkes process still admits
closed-form posteriors for the background rate; the branching ratio requires
MCMC or variational inference.}

\textbf{The windowed model discards information.} Using only the last 14
days is the simplest cure for non-stationarity, but it throws away all
historical context. A hierarchical model with a changepoint prior---or a
particle filter over a regime-switching Poisson---would allow the model
to both detect regime changes and retain memory within each regime.

\textbf{No real validation.} Because we use synthetic data, we cannot
compare the model's crowding forecasts against actual ICU outcomes.
Validating on a real MIMIC-IV cohort---with proper IRB approval and
de-identification---is the necessary next step before any clinical application.

Despite these limitations, BUICU demonstrates something important: a
probabilistic forecast that is \emph{honest about what it does not know}
is more useful than a deterministic prediction that pretends certainty.
Every number in BUICU carries a credible interval. The model knows when it
is surprised. And when its assumptions break down, it says so---widening
the interval rather than projecting false confidence.

% ─────────────────────────────────────────────────────────────────────
\subsection*{Ethical Reflection}

A model that informs staffing decisions in a critical-care setting carries
real ethical weight. Three safeguards are built into BUICU from the start.

First, \textbf{synthetic data only}. Real ICU records contain protected
health information. Using a synthetic generator---calibrated to published
MIMIC-IV statistics but containing no real patient data---eliminates the
privacy risk while preserving statistical realism.

Second, \textbf{uncertainty is non-negotiable}. Every forecast is a
distribution, never a point estimate. The website never says ``there will
be 12 admissions tomorrow''; it says ``we expect between 8 and 16 with
95\% probability''. Overconfident predictions in clinical settings have
a name: malpractice.

Third, \textbf{the model discloses its own failure modes}. The five failure
modes are not an appendix; they are front-page content on the website,
complete with detection evidence, severity ratings, and a combined
uncertainty-widening factor. A model that cannot articulate how it fails
should not be trusted.

The feedback-loop failure mode deserves special emphasis. If staff
use crowding forecasts to pre-emptively divert patients, the observed
arrival rate reflects the \emph{managed} system. The model then learns the
managed rate and lowers its future estimates---reducing perceived urgency
and potentially triggering a dangerous relaxation of interventions. This is
Goodhart's Law in its most literal form. BUICU must never be deployed for
autonomous decision-making without explicit human oversight and a
documented protocol for distinguishing managed from unmanaged arrival data.

\end{document}
